---
title: "HW4"
output: html_document
---
### Name:Wei Chen
### ID:5207679
Exercise 4.19
Let X~1~,X~2~,...,X~20~ be a random sample of size n=20 from an N~6~($\mu$, $\Sigma$) population.
(a)The distribution of (X~1~-$\mu$)'$\Sigma$(X~1~ $\mu$)      
(b)The distribution of $\bar{X}$ and $\sqrt{n}(\bar{X}-\mu)$    
(c)The distribution of (n-1)S   

Solution:   
(a)   
Let Z=$\sigma^{-1/2}(X_{1}-\mu)$,then  (X~1~-$\mu$)'$\Sigma$(X~1~ $\mu$)=Z'Z.  
Since X~1~ ~N ~6~($\mu$, $\Sigma$)and Z~N(0,1), Therefore, the squred standard normal distribution is $\chi^{2}_{1}$ distribution with 1 degree of freedom.   
(X~1~-$\mu$)'$\Sigma$(X~1~ $\mu$)~ $\chi^{2}_{1}$   
(b) 
$\bar{X}$ =$\frac{1}{n}\sum_{i=1}^{20}X_{i}$ ~ N~6~($\mu$,$\Sigma$).         $\sqrt{n}(\bar{X}-\mu)$ ~ N~6~(0,$\frac{\Sigma}{\sqrt{20}}$)                    
(c) (n-1)S follows the Wishart distribution with n-1 degrees of freedom. (n-1)S~ W~19~($\Sigma$)

Exercise 4.20
Specify the B(19S)B' in each case     
(a) B=\[ \left( \begin{array}{ccc}
1 & -1/2 & -1/2 & 0 & 0 & 0 \\
0 & 0 & 0 &-1/2 &-1/2 &-1/2 \\
\end{array} \right)\] 
(b)B=\[ \left( \begin{array}{ccc}
1 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 1 & 0 & 0 & 0 \\
\end{array} \right)\] 

Solution:
Let X=B'(19S)B        
(X~1~-$\mu$)'$\Sigma$(X~1~ $\mu$)~ $\chi^{2}_{1}$   
$\bar{X}$ ~ N~6~($\mu$,$\Sigma$)    
$\sqrt{n}(\bar{X}-\mu)$ ~ N~6~(0,$\frac{\Sigma}{\sqrt{20}}$)        
(20-1)S~ W~19~($\Sigma$)           
(a)   

```{r}
B<- matrix(c(1, -0.5,-0.5, 0, 0,0,0,0,0,-0.5,-0.5,-0.5),nrow=2, ncol=6 )
X<-B%*%cov(B)%*%t(B)
mu<-colMeans(B)
sigma<- cov(B)
print(mu)
print(sigma)
```

(b)   
```{r}
B<- matrix(c(1, 0,0, 0, 0,0,0,0,1,0,0,0),nrow=2, ncol=6 )
X<-B%*%cov(B)%*%t(B)
mu<-colMeans(B)
sigma<- cov(B)
print(mu)
print(sigma)
```

## Problem2

1Proof:   
Design a special matrix P to transfrom X=[X~1~ X~2~ ... X~n~]'  to Z so that the first entry of Z is related to $\bar{X}$ and the rest is related to S^2^.
Let P be orthogonal matrix and the first enry P~1~ =[$\frac{1}{\sqrt{n}}$ $\frac{1}{\sqrt{n}}$... $\frac{1}{\sqrt{n}}$]'.      
$\because$ P matrix is orthognal, P'P=I    
$\therefore$ Z'Z=X'P'PX= X'X    
$\because$ P~1~ =[$\frac{1}{\sqrt{n}}$ $\frac{1}{\sqrt{n}}$... $\frac{1}{\sqrt{n}}$]'   
$\therefore$ Z~1~=P'~1~ X=$\sqrt{n}$$\bar{X}$      
$\therefore$ $\bar{X}$=$\frac{1}{\sqrt{n}}$Z~1~              
(n-1)S^2^=X'X- n$\bar{X}^{2}$ =Z'Z -n$(\frac{1}{\sqrt{n}}Z_{1})^{2}$ =Z'Z-Z~1~^2^= $\sum_{i=2}^{n}$Z~i~^2^       
cov(Z~1~, Z~i~)=0, i=2,3,...n       
$\bar{X}$ and S^2^ are independent.

      
2   
(a)
Y=[Y~n1~ Y~n2~]'= AZ    
The first row has n1 1 and second row has n2 1.   
\[A={\begin{pmatrix}
1 & 1 & \cdots & 1_{n1} & 0 & ...& 0 \\
1 & 1 & \cdots & 1 & 1 & ...& 1 \\
\end{pmatrix}}_{2*n2} \]
Z=[X~1~ X~2~ ... X~n2~]'    
Since X~1~ X~2~ ... X~n2~ are independent and follows N($\mu$,$\Sigma$)   
Z~N($\mu_{z}$, $\Sigma_{z}$)    
$\mu_{z}$=[$\mu$', $\mu$',...,$\mu$']'~n2~   
$\Sigma_{z}$=$\Sigma$ I~n2~   
AZ~N(A$\mu_{z}$, A$\Sigma$ A')        
A$\mu_{z}$=[n~1~$\mu$'  n~2~$\mu$']'   

\[A\Sigma A' ={\begin{pmatrix}
n_{1}\Sigma & n_{1}\Sigma \\
n_{1}\Sigma  & n_{2}\Sigma  \\
\end{pmatrix}} \]   
(b)   
Y~1|2~ ~ N~q~($\mu_{1|2}$,$\Sigma_{1|2}$)    
$\mu_{1|2}$=$\mu_{1}$+$\Sigma_{12}$    $\Sigma_{22}^{-1}$(y-$\mu_{2}$)=[$\mu$'...$\mu$']'~n1~+$\frac{n1}{n2}$ (y-[$\mu$'...$\mu$']'~n1~)        
$\Sigma_{1|2}$=$\Sigma_{11}$-$\Sigma_{12}$$\Sigma_{22}^{-1}$$\Sigma_{21}$ = $n_{1}$$\Sigma$-$\frac{n_{1}^{2}}{n_{2}}$$\Sigma$
