---
title: "Assignment1"
output: html_document
---
### *Wei Chen 5207679*
### *Jan/29/2016* 
##Exercise 1.4:

(a)	plot the scatter diagram with marginal boxplots for sales and profits variables
(b) compute $\bar{\mathbb{x}}$~1~ ,$\bar{\mathbb{x}}$~2~, s~11~, s~22~, s~12~ and r~12~ 

##Solution:
(a)First, I need to install packages "alr4" and then library it.
```{r}
install.packages("alr4",repos = 'http://cran.us.r-project.org')
library(alr4)

```

Now I import data:
````{r}
p1_4 <-read.table("http://users.stat.umn.edu/~guxxx192/courses/wichern_data/p1-4.dat")
```

I add the column and plot the scatter diagram with marginal boxplots
```{r}
colnames(p1_4) <- c("sales","profits","assets")
scatterplot(sales ~ profits, data= p1_4, smoother= F, reg.line= F)
```

（b）In order to calculate sales and profits variablrs, I first remove other variables and create a new data set p1_4_new
```{r}
p1_4_new <-p1_4[,c("sales","profits")]
```
And then print the mean of sales and profits
```{r}
print(colMeans(p1_4_new))
```
Now I calculate the variance matrix and print S_1_1, S_2_2, S_1_2
```{r}
S=var(p1_4_new)
c(S[1,1] ,S[2,2],S[1,2])
```
Calculate the correlation and print the r_1_2
```{r}
C<- cor(p1_4_new)
C[1,2]
```
It shows that correlation 0.68 between sales and profits is positive which proximately means positive linear association. They change almost in same direction that the more sales is, the more the profits is.

##Exercise 1.5
Use the data in Exercise 1.4
(a) plot the scatter diagrams and boxplots for (x~2~, x~3~), (x~1~, x~3~). Comment on the patterns
(b) Compute the $\bar{\mathbb{x}}$, S~n~, and R arrays for (x~1~, x~2~, x~3~)

##Solution:
(a) Plot (x~2~, x~3~) and find no explicit pattern. The same amount of assets corresponds to different profits. 
```{r}
scatterplot(assets~profits, data= p1_4, smoother=F, reg.line= F)
```

plot (x~1~, x~3~) and find that assets go down as the sales become larger 
```{r}
scatterplot(assets~sales, data =p1_4, smoother= F, reg.line= F)
```

(b) array X_mean=
```{r}
colMeans(p1_4)
```
variance matrix S_n =
```{r}
var(p1_4)
```
correlation matrix R_n =
```{r}
cor(p1_4)
```
##Exercise 1.15
Some of the 98 measurements described in Section 1.2 are listed in Table 1.7. The data consists of average ratings over the course of treatment for patients undergoing reditherapy. Variables measured include x~1~(number of symptoms); x~2~(amount of activity); x~3~(amount of sleep,on a 1-5 scale); x~4~(amount of food consumed, on a 1-3 scale); x~5~(appetite, on a 1-5 scale); x~6~(skin reaction, on a 0-3 scale).
(a) Construct the two-dimensional scatterplot for variables x~2~ and x~3 and the margnal boxplots. Do there appear to be any errors in the x~3~ data?
(b)Compute the $\bar{\mathbb{x}}$, S~n~, and R arrays. Interpret the pairwise correlations.

##Solution:
(a)First I import data from http://users.stat.umn.edu/~guxxx192/courses/wichern_data/t1-7.dat
```{r}
t1_7<- read.table("http://users.stat.umn.edu/~guxxx192/courses/wichern_data/t1-7.dat")
```
I add column names and then conduct scatter plot with boxplots.x~3~ should be included in 1-5 scale, but there is a few points lower than 1 and no points higher than 4. So there may be some mistake.
```{r}
colnames(t1_7)<-c("x_1","x_2","x_3","x_4","x_5","x_6")
scatterplot(x_3~x_2, data= t1_7, smoother=F, reg.line=F)
```

(b)Calculate the $\bar{\mathbb{x}}$ :
```{r}
colMeans(t1_7)
```
Calculate the S~n~:
```{r}
var(t1_7)
```
Calculate the R:
```{r}
cor(t1_7)
```
The diagnal is the correlation between same varaibles and it is always 1.The R matrix is symmetric.
It implies little change direction tendency for pair(amount of activity,skin reaction), (number of symptoms,skin reaction),(amount of food consumed,skin reaction),(appetite,skin reaction). It also implies a same change direction tendency for other pairs, such as (amount of apetite,amount of food consumed),(appetite, number of symptoms)and (amount of activity, number of symptoms).
##Exercise 1.16
At the start of a study to determin thether exercise or dietary supplements would slow bone loss in older women, an investigateor measured the mineral content of bones by photon absorptiometry. Measurement were recorded for three bones on the dominant and nondominant sides and are shown in Table 1.8.
Compute the $\bar{\mathbb{x}}$, S~n~, and R arrays. Interpret the pairwise correlations.
##Solution:
First import data from http://users.stat.umn.edu/~guxxx192/courses/wichern_data/t1-8.dat
```{r}
t1_8<- read.table("http://users.stat.umn.edu/~guxxx192/courses/wichern_data/t1-8.dat")
```
Add the column names and calculate the mean $\bar{\mathbb{x}}$=
```{r}
colnames(t1_8)<- c( "Dominant radius","Radius", "Dominant humerus", "Humerus","Dominant ulna","Ulna")
colMeans(t1_8)
```
Calculate the variance S~n~ =
```{r}
var(t1_8)
```
Calculate the correlation R=
```{r}
cor(t1_8)
```
Most of the correlation between variables are significant. The smallest is 0.44 for (Ulna, Dominant humerus) and the largest is 0.89 for (Dominant humerous, Humerous).
##Exercise 1.18
Convert the national track records for women in Table 1.9 to speeds measured in meters per second. For example, the record spped for the 100-m dash for Argentinian women is 100m/11.57 sec=8.643 m/sec. Notice that the records for the 800-m, 1500-m, 3000-m and marathon runs are measured in minutes. The marathon is 26.2 miles, or 42,195 meters long.Compute the $\bar{\mathbb{x}}$, S~n~, and R arrays. Notice the magnitudes of the correlation coefficient as you go from the shorter (100m) to the longer(marathon) running distances. Interpret these pairwise correltaions. Compare your results with the results you obtained in Exercise 1.17
##Solution
First, I import data from http://users.stat.umn.edu/~guxxx192/courses/wichern_data/t1-9.dat with argument sep="\t" that prevents R from treating the spaces in "KOR, S" and "KOR, N" as separators.
```{r}
t1_9<- read.table("http://users.stat.umn.edu/~guxxx192/courses/wichern_data/t1-9.dat", sep="\t")
```
Add the column names and after reomve the "country column", calculate the number of rows in table 1.9 in order to create the convert matrix to change the measurement to speed.
```{r}
colnames(t1_9)<-c("Country","100m","200m","400m","800m","1500m","3000m","Marathon")
t1_9_time<- t1_9[,-c(1)]
nrow(t1_9)
```
Then create the convert matrix to change the measurement to speed as matirx, t1_9_speed.
```{r}
convert <-c(rep(100,54),rep(200,54),rep(400,54),rep(800/60,54),rep(1500/60,54),rep(3000/60,54),rep(42195/60,54))
t1_9_speed<- convert/t1_9_time 
t1_9_speed
```
$\bar{\mathbb{x}}$_speed=
```{r}
colMeans(t1_9_speed)
```
S~n~_speed=
```{r}
var(t1_9_speed)
```
Calculate Array R_speed, and find that magnitude of correlation goes down as we go from the shorter (100m) to the longer(marathon) running distances. But all variables have positive correlation which means the country has fast speed record in one distance tends to have a fast speed record in other distance.
R_speed=
```{r}
cor(t1_9_speed)
```

Compare with the Exercise 1.17 measured in time and find that although they have subtle difference, they have same tendency.
R_time=
```{r}
cor(t1_9_time)
```

##Exercise 1.19
Create the scatter plot and boxplot displays of Figure 1.5 for (a) the mineral-content data in Table 1.8 and (b) the national-track-records data in Table 1.9.

##Solution:
For all columns in Table1.8, I plot them with boxplots
```{r}
scatterplotMatrix(t1_8[,1:ncol(t1_8)], diagonal="boxplot", smoother=F,reg.line=F)
```

For all columns in Table1.9, I plot them with boxplots
```{r}
scatterplotMatrix(t1_9_time[,1:ncol(t1_9_time)],diagonal="boxplot", smoother=F, reg.line=F)
```