---
title: "Assign2"
output: html_document
---
### *Wei Chen 5207679*
### *Feb/5/2016* 
## Problem 1
### 2.4(p.103)
(a)
∵ AA^-1^=I  
∴ (AA^-1^)'=I'=I  
∴ (A^-1^)'A'=I  
∴(A^-1^)'A'(A')^-1^=(A')^-1^  
∴(A^-1^)'I=(A')^-1^   
∴(A^-1^)'=(A')^-1^  
(b)
∵ (B^-1^A^-1^)AB=B^-1^(A^-1^A)B=B^-1^B=I  
∴ (B^-1^A^-1^)AB(AB)^-1^=(AB)^-1^ 
∴ (B^-1^A^-1^)=(AB)^-1^ 

### 2.10(P.104)
A^-1^=
```{r, echo=FALSE}
A<- matrix(c(4, 4.001, 4.001, 4.002), nrow=2, ncol=2)
B<-matrix(c(4, 4.001, 4.001, 4.002001), nrow=2, ncol=2)
A_inv<-solve(A)
B_inv<-solve(B)
```
```{r}
A_inv
```
B^-1^=
```{r}
B_inv
```
To see whether A_inv is -3 times of B_inv
```{r}
(-3)*B_inv 
```
So -3*B_inv is euqal to A_inv

##2.11(p.104)
By the definition 2A.24,  
$|A_{p\times p}|$ =a~11~$\times$$|A_{p-1\times p-1}|$=a~11~$\times$a~22~ $\times$ $|A_{p-2\times p-2}|$ =a~11~$\times$a~22~$\times$a~33~...a~(p-1)(p-1)~$\times$ $|A_{1\times 1}|$=a~11~ $\times$a~22~$\times$a~33~...a~(p-1)(p-1)~$\times$a~pp~

###2.13(p.104)
Since Q is orthogonal matrix  
∴QQ'=I  
∴|QQ'|=|I|  
∴|Q||Q'|=|I|=1  
∵ Q is a square matrix  
∴ |Q|=|Q'|  
∴ |Q||Q'|=|Q|^2^=1  
∴ |Q|=+1 or -1  

###2.14(p.104)
Let $\lambda$ be any eigenvalue of A, then $|A-\lambda I|$ =0   
So $|Q'||A-\lambda I||Q|$ =0    
∵ $|A||B|=|AB|$     
∴ $|Q'||A-\lambda I||Q|$ =$|Q' A-\lambda Q-\lambda Q'Q|$=0  
∵ Q'Q=I since Q is orthogonal   
∴ $|Q' AQ-\lambda I|$=0   
∴ Q' AQ has all the eigenvalues that A has    
∵ Q' AQ is $p \times p$ dimension which is the same to A    
∴ Q' AQ can not have more eigenvaue than A    
∴ Q' AQ and A have the same eigenvalues if Q is orthogonal    

##2.16(p.104)
Let x be a $p \times 1$ vector, then x'A'Ax will be a scalar.   
Let y=Ax, y will be $n \times 1$ vector.    
x'A'Ax= y'y=y~1~^2^+y~2~^2^+...+y~n~^2^ >=0 for all x vector.   
Thus A'A is neccessarily nonnegative definite.

##2.17
Let $\lambda$ be the eigenvalue of A  
According to the eigenvalue deinition, Ae=$lambda$ e, where e is eigenvector.  
Multipied by e', we get e'Ae=e'$lambda$ e  
Since A is positive definite  
e'Ae>0 for e$\neq$ 0   
e'Ae=e'$lambda$ e=$lambda$>0, since eigenvector is non-zero  

##2.18
$4x^{1}_{2} +3x^{2}_{2}=2 \sqrt x_{1}x_{2}= x'Ax$, where x=[x~1~ x~2~]',   
A= \[ \left( \begin{array}{ccc}
4 & - \sqrt 2 \\
-\sqrt 2& 3 \end{array} \right)\]
axis x~j~=c $\lambda_{j}^{ -1/2}$  e~j~

```{r echo= FALSE}
A<- matrix(c(4, -sqrt(2), -sqrt(2), 3), nrow=2, ncol=2)
```
A=
```{r}
A
```
 eigenvalues $\lambda_{j}$ = 
```{r}
eigVal<-eigen(A)$values
eigVal
```
eigenvectors e~j~=
```{r}
eigVec<-eigen(A)$vectors
eigVec
```
for c^2^=1,   
minor axis is x~1~= 1 /$\sqrt 5$=$\sqrt 5$/5  e~1~, e~1~=[-0.81, 0.57]'  
major axis is x~2~=1/ $\sqrt 2$=$\sqrt 2$/2  e~2~, e~2~=[0.57,-0.81]'  
for c^2^=4  
minor axis is x~1~= 2 /$\sqrt 5$=2$\sqrt 5$/5  e~1~,  e~1~=[-0.81, 0.57]'  
major axis is x~2~=2/ $\sqrt 2$=$\sqrt 2$  e~2~,  e~2~=[0.57,-0.81]'   
If we sketch the ellipses, we get 
```{r}
library(car)
ellipse(c(0, 0), shape=A, radius=0.98, col="red", lty=2,add=FALSE)
ellipse(c(0, 0), shape=A/4, radius=0.98, col="red", lty=2)
arrows(0, 0, sqrt(eigVal[1]) * eigVec[ , 1][1], sqrt(eigVal[1]) * eigVec[ , 1][2])
arrows(0, 0, sqrt(eigVal[2]) * eigVec[ , 2][1], sqrt(eigVal[2]) * eigVec[ , 2][2])
```


The two ellipse have same axes direction and length increase as c^2^ increase

###2.38
```{r}
library(car)
A<-matrix(c(13,-4,2,-4,13,-2,2,-2,10),3,3)
eigen(A,only.values=FALSE)
```
According to Extended Cauchy-Schwarz Inequality
The maximum is 18, while the mnimum is 9

## Problem 2
### 1
(1)In order to prove:
\[
\begin{pmatrix} 
A & O  \\
C & B \\
\end{pmatrix}^{-1}=
\begin{pmatrix} 
A_{-1} & O  \\
-B^{-1}CA^{-1} & B^{-1} \\
\end{pmatrix} \]
We need to prove:
\[
\begin{pmatrix} 
A_{-1} & O  \\
-B^{-1}CA^{-1} & B^{-1} \\
\end{pmatrix} \times
\begin{pmatrix} 
A & O  \\
C & B \\
\end{pmatrix}= 
\begin{pmatrix}
AA^{-1} & O \\
-B^{-1}CA^{-1}A+B^{-1}C & B^{-1}B \\
\end{pmatrix} =
\begin{pmatrix}
I & O \\
O & I\\
\end{pmatrix}
\]
(2)And in order to prove 
\[
\begin{pmatrix}
A & D \\
O & B \\
\end{pmatrix}^{-1}=
\begin{pmatrix}
A^{-1} & -A^{-1}DB^{-1} \\
O & B^{-1}\\
\end{pmatrix}
\]
We need to prove:
\[
\begin{pmatrix}
A^{-1} & -A^{-1}DB^{-1} \\
O & B^{-1}\\
\end{pmatrix} \times
\begin{pmatrix}
A & D \\
O & B \\
\end{pmatrix}= 
\begin{pmatrix}
A^{-1}A & A^{-1}D-A^{-1}DB^{-1}B \\
O B^{-1}B\\
\end{pmatrix}=
\begin{pmatrix}
I & O \\
O & I  \\
\end{pmatrix}
\]

### 2
Let T=A+CBD   
In order to prove T^-1^=A^-1^-A^-1^CB(B+BDA^-1^CB)^-1^BDA^-1^  
We need to prove (A^-1^-A^-1^CB(B+BDA^-1^CB)^-1^BDA^-1^)T=I  
In other words,     
(A^-1^-A^-1^CB(B+BDA^-1^CB)^-1^BDA^-1^)(A+CBD)   
=A^-1^-A^-1^CB(B+BDA^-1^CB)^-1^BDA^-1^A+A^-1^CBD-A^-1^CB(B+BDA^-1^CB)^-1^BDA^-1^CBD  
=I-(A^-1^CB(B+BDA^-1^CB)^-1^)(BD-(B+BDA^-1^CB)D+BDA^-1^CBD)       
=I-(A^-1^CB(B+BDA^-1^CB)^-1^)O  
=I    
In particular, when q=1, B=1, C=c,D=d'  
We want to show that (A+cd')^-1^ =A^-1^ -${ {A^{-1}cd'A^{-1} }\over {1+d'A~{-1}c }}$  
substitue into T,         
T^-1^=(A+cd')^-1^=A^-1^-A^-1^c(1+d'A^-1^c)^-1^d'A^-1^     
Since d'A^-1^c is scalar    
T^-1^=A^-1^-A^-1^cd'A^-1^ /(1+d'A^-1^c)

### 3
For A=\[
\begin{pmatrix}
A_{11} & A_{12} \\
A_{21} & A_{22} \\
\end{pmatrix}
\]
In order to state clearly in the deduction process, we use B to denote A^-1^=B=\[
\begin{pmatrix}
B_{11} & B_{12}\\
B_{21} & B_{22}\\
\end{pmatrix}=
\begin{pmatrix}
A_{11}^{-1}+A_{11}^{-1}A_{12}A_{22·1}^{-1}A_{21}A_{11}^{-1} & -A_{11}^{-1}A_{12}A_{22·1}^{-1}\\
-A_{22·1}^{-1}A_{21}A_{11}^{-1} & A_{22·1}^{-1}
\end{pmatrix}
\]
where A~22·1~=A~22~ - A~21~A~11~^-1^A~12~   
We can prove for B~11~, B~12~, B~21~, B~22~   
B~11~A~11~+B~12~A~21~= A~11~^-1^A~11~ + A~11~^-1^A~12~A~22·1~^-1^A~21~A~11~^-1^A~11~-A~11~^-1^A~12~A~22·1~^-1^A~21~    
=I+O    
=I    
B~11~A~12~+B~12~A~22~=A~11~^-1^A~12~+A~11~^-1^A~12~A~22·1~^-1^A~21~A~11~^-1^A~12~ - A~11~^-1^A~12~A~22·1~^-1^A~21~    
= A~11~^-1^A~12~ + A~11~^-1^A~12~A~22`1~^-1^(A~21~A~11~^-1^A~12~-A~22~)    
= A~11~^-1^A~12~ + A~11~^-1^A~12~A~22·1~(-A~22·1~)    
=O    
B~21~A~11~+B~22~A~21~ = -A~21~^-1^A~21~ + A~22·1~^-1^ A~21~ =  O    
B~21~A~12~+B~22~A~22~ = -A~22·1~A~21~A~11~^-1^A~12~ + A~22·1^-1^~A~22~    
= A~22·1~^-1^(A~22~-A~21~A~11~^-1^A~12~)    
=A~22·1~^-1^A~22·1~^-1^
= I   
Thus, BA=I,     
A^-1^=B=\[
\begin{pmatrix}
B_{11} & B_{12}\\
B_{21} & B_{22}\\
\end{pmatrix}=
\begin{pmatrix}
A_{11}^{-1}+A_{11}^{-1}A_{12}A_{22·1}^{-1}A_{21}A_{11}^{-1} & -A_{11}^{-1}A_{12}A_{22·1}^{-1}\\
-A_{22·1}^{-1}A_{21}A_{11}^{-1} & A_{22·1}^{-1}
\end{pmatrix}
\]
