---
title: "ID: 5207679"
output: html_document
---
##Exercise 4.2 (c)
the constant density contour that contains 50% of the probability
```{r}
library(car)
sigma<- matrix(c(2,sqrt(0.5),sqrt(0.5),1),nrow=2,ncol=2)
ellipse(center=c(0,2), shape=sigma, radius=sqrt(qchisq(0.5, df=2)),add=FALSE)
text(1,2,label="50%")
```

## Exercise 4.29 (a)
Caluculate the statistical distances of 42 points per variable x5 and x6. 
```{r}
t1_5 <- read.table("http://users.stat.umn.edu/~guxxx192/courses/wichern_data/t1-5.dat")
t1_5new <- t1_5[,c(5:6)]
Y=as.matrix(t1_5new)
D<- mahalanobis(Y, center=colMeans(Y), cov=cov(Y))
print(D)
```
##Exercise 4.29 (b) 
The center of ellipse is 
```{r}
colMeans(Y)
```
plot the 50% probability contour and count the proportion of observations that fall into within the contour. 
```{r}
X <-t(Y)
x6<-X[2,]
x5<-X[1,]
plot(x6~x5)
ellipse(center=c(10,9.4), shape=cov(Y), radius=sqrt(qchisq(0.5, df=2)))
text(14,10,label="50%")
```

Almost 26 of 42 observations fall within the contour, the proportion is 0.62. 
